# Objective
Here I will analyse the T and B cell clone datasets inferred from bulk RNA-Seq using miXCR. The main objective of this analysis is finding if the TCR & BCR clonotypes that I see in a tumour region can be observed in other tumour regions, even in distal locations such as the metastases. 

# ANALYSIS
## 1. Set-up and load annotation and clone data

```{r setup, include=FALSE}
rm(list = ls(all = TRUE))
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(results = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_knit$set(root.dir = file.path(here::here(), "new_samples")) # change accordingly
knitr::opts_chunk$set(fig.path = file.path(here::here(), "new_samples", "analyses", "miXCR"))
```

```{r paths}
# PATHS
BASE <- here::here()
META_DIR = file.path(BASE, "data", "meta")
OUT_DIR = file.path(BASE, "analysis", "outputs")
FIG_DIR = file.path(BASE, "analysis", "figures")
RAW_DIR = file.path(BASE, "data", "raw")
ANNOTATION_PATH <- file.path(META_DIR, "tx_annotation.tsv")
TCR_CLONES_PATH <- file.path(RAW_DIR, "tumour_TCR_clones.RDS")
TCR_CLONES_NORMALS_PATH <- file.path(RAW_DIR, "normal_TCR_clones.RDS")
BCR_CLONES_PATH <- file.path(RAW_DIR, "tumour_BCR_clones.RDS")
BCR_CLONES_NORMALS_PATH <- file.path(RAW_DIR, "normal_BCR_clones.RDS")
```

```{r}
library(tidyverse)
library(immunarch)
library(pheatmap)
library(lemon)
library(ggpmisc)
library(ggthemes)

source(file.path(BASE, "src", "plotting_theme.R"))
source(file.path(BASE, "src", "plotting_functions.R"))
```

```{r}
annotation <- read_delim(ANNOTATION_PATH, delim = "\t")
TCRClonotypes <- readRDS(TCR_CLONES_PATH)
BCRClonotypes <- readRDS(BCR_CLONES_PATH)
normalsTCRClonotypes <- readRDS(TCR_CLONES_NORMALS_PATH)
normalsBCRClonotypes <- readRDS(BCR_CLONES_NORMALS_PATH)


# exclude all the samples that corresponds to patients with one single sample analysed
patients <- str_extract(names(TCRClonotypes$data), "K\\d{3}")
filterPatients <- names(table(patients) > 1)
keepSamples <- which(patients %in% filterPatients)
TCRClonotypes <- list(data = TCRClonotypes$data[keepSamples], meta = TCRClonotypes$meta[keepSamples, ])
names(TCRClonotypes$data) <- str_remove(names(TCRClonotypes$data), "R_") %>% str_remove("G_") %>% str_remove(".clonotypes.TCRA_TCRB") %>% 
  str_remove("_r\\d") %>% str_replace("K328R19", "K328-R19")
TCRClonotypes$meta$Sample <- names(TCRClonotypes$data)

BCRClonotypes <- list(data = BCRClonotypes$data[keepSamples], meta = BCRClonotypes$meta[keepSamples, ])
names(BCRClonotypes$data) <- str_remove(names(TCRClonotypes$data), "R_") %>% str_remove("G_") %>% str_remove(".clonotypes.IGK_IGL") %>% 
  str_remove("_r\\d") %>% str_replace("K328R19", "K328-R19")
BCRClonotypes$meta$Sample <- names(BCRClonotypes$data)
```

```{r functions}
classify_clones <- function(samples, clonotypes){
  # find all the possible pairs where to analyse the matches
  pairs <- combn(samples, 2)
  lClones <- lapply(1:ncol(pairs), function(indexPair){find_match_clones(pairs[,indexPair][1], pairs[,indexPair][2], clonotypes)})
  names(lClones) <- paste0(pairs[1,], "|", pairs[2, ])
  return(lClones)
}


# function to get subset of the dataframe in both samples with the 
find_match_clones <- function(sampleID1, sampleID2, clonotypes){
  clones_sample1 <- clonotypes$data[[sampleID1]]
  clones_sample2 <- clonotypes$data[[sampleID2]]
  common_clones <- dplyr::intersect(clones_sample1$CDR3.nt, clones_sample2$CDR3.nt)
  shared_clones_sample1 <- clones_sample1[clones_sample1$CDR3.nt %in% common_clones, c("Proportion", "CDR3.nt")]
  colnames(shared_clones_sample1) <- c(paste0(sampleID1, "_proportion"), "CDR3.nt")
  shared_clones_sample2 <- clones_sample2[clones_sample2$CDR3.nt %in% common_clones, c("Proportion", "CDR3.nt")]
  colnames(shared_clones_sample2) <- c(paste0(sampleID2, "_proportion"), "CDR3.nt")
  private_clones_sample1 <- clones_sample1[! clones_sample1$CDR3.nt %in% common_clones, c("Proportion", "CDR3.nt")]
  colnames(private_clones_sample1) <- c("Proportion", "CDR3.nt")
  private_clones_sample2 <- clones_sample2[! clones_sample2$CDR3.nt %in% common_clones, c("Proportion", "CDR3.nt")]
  colnames(private_clones_sample2) <- c("Proportion", "CDR3.nt")
  return(list(shared_clones = merge(shared_clones_sample1, shared_clones_sample2, by = "CDR3.nt"), 
              private_clones = rbind(private_clones_sample1, private_clones_sample2)))
}

is_same_pt <- function(sample_pair_names){
  str_extract_all(sample_pair_names, "K[^_-]*") %>% # what goes after K and before _ or - is patient ID
    {unlist(lapply(., function(thisPair) length(unique(thisPair)) == 1))} # check if the values are the same with unique(thisPair) == 1
}
```

## 2. Analysis of conserved T and B cells in different tumour regions

### TCR
```{r TCR}
# compute the total number of matching clones
sharedTCRMat <- immunarch::repOverlap(TCRClonotypes$data, .method = "public")
# arrange samples of the same patient together
sharedTCRMat <- sharedTCRMat[sort(rownames(sharedTCRMat)), sort(colnames(sharedTCRMat))]
png(filename = file.path(FIG_DIR, "sharedTCR_heatmap.png"), width = 80, height = 80, units = "mm", res = 400)
pheatmap(sharedTCRMat, cluster_rows = FALSE, cluster_cols = FALSE, show_colnames = FALSE, show_rownames = FALSE, color = hcl.colors(4, "Blues"), breaks = c(0, 1, 15, max(sharedTCRMat, na.rm = TRUE)), legend = FALSE)
dev.off()
```

We can clearly see that the TCR clones that have the same sequence are usually found in the same patient. Let's corroborate this observation

```{r}
lTCRClones <- classify_clones(names(TCRClonotypes$data), TCRClonotypes)
unspecificTCRShared <- c()
numberUnspecificPairs <- 0
specificTCRShared <- c()
numberSpecificPairs <- 0

for (thisPair in 1:length(lTCRClones)){
  if (str_detect(names(lTCRClones)[thisPair], "M|LN|T")){
    next()
  }
  if (nrow(lTCRClones[[thisPair]]$shared_clones)>0){
    if(! is_same_pt(names(lTCRClones)[thisPair])){
      unspecificTCRShared <- c(unspecificTCRShared, lTCRClones[[thisPair]]$shared_clones$CDR3.nt)
      numberUnspecificPairs <- numberUnspecificPairs + 1
    } else {
      specificTCRShared <- c(specificTCRShared, lTCRClones[[thisPair]]$shared_clones$CDR3.nt)
      numberSpecificPairs <- numberSpecificPairs + 1
    }
  }
}
```

```{r}
nSpecificPairsPrimPrim_tcr = numberSpecificPairs
nPrimPrim_same_pat_pairs = sum(sapply(1:length(lTCRClones), function (thisPair){
  is_same_pt(names(lTCRClones[thisPair])) & 
    !str_detect(names(lTCRClones)[thisPair], "M|LN|T")
}))

nUnspecificPairsPrimPrim_tcr = numberUnspecificPairs
nPrimPrim_diff_pat_pairs = sum(sapply(1:length(lTCRClones), function (thisPair){
  !is_same_pt(names(lTCRClones[thisPair])) & 
    !str_detect(names(lTCRClones)[thisPair], "M|LN|T")
}))
```

1063 different TCR clones overlapping in 204 / 262 pairs of primary sample pairs from the same patient that share at least one TCR clone. 115 different TCR clones overlapping in 72 / 19838 pairs of primary sample pairs that share at least one TCR clone. 


```{r}
unspecificTCRSharedMets <- c()
numberUnspecificPairsMets <- 0
specificTCRSharedMets <- c()
numberSpecificPairsMets <- 0
props = data.frame(s1 = c(), s2 = c(), prop_s1 = c(), prop_s2 = c(), clone = c())
for (thisPair in 1:length(lTCRClones)){
  if (str_detect(names(lTCRClones)[thisPair], "M|LN|T") & str_detect(names(lTCRClones)[thisPair], "R")){
    if (nrow(lTCRClones[[thisPair]]$shared_clones)>0){
      if(! is_same_pt(names(lTCRClones)[thisPair])){
        unspecificTCRSharedMets <- c(unspecificTCRSharedMets, lTCRClones[[thisPair]]$shared_clones$CDR3.nt)
        numberUnspecificPairsMets <- numberUnspecificPairsMets + 1
      } else {
        specificTCRSharedMets <- c(specificTCRSharedMets, lTCRClones[[thisPair]]$shared_clones$CDR3.nt)
        df = lTCRClones[[thisPair]]$shared_clones
        s1 = colnames(df)[ncol(df) - 1]
        s2 = colnames(df)[ncol(df)]
        prop_s1 = as_vector(df[,ncol(df) - 1])
        prop_s2 = as_vector(df[,ncol(df)])
        clone = df$CDR3.nt
        props = rbind(props, data.frame(s1 = s1, s2 = s2, prop_s1 = prop_s1, prop_s2 = prop_s2, clone = clone))
        numberSpecificPairsMets <- numberSpecificPairsMets + 1
      }
    }
  }
}
```


```{r}
nSpecificPairsPrimMet_tcr = numberSpecificPairsMets
nPrimMet_same_pat_pairs = sum(sapply(1:length(lTCRClones), function (thisPair){
  is_same_pt(names(lTCRClones[thisPair])) & 
    str_detect(names(lTCRClones)[thisPair], "M|LN|T") &
    str_detect(names(lTCRClones)[thisPair], "R")
}))

nUnspecificPairsPrimMet_tcr = numberUnspecificPairsMets
nPrimMet_diff_pat_pairs = sum(sapply(1:length(lTCRClones), function (thisPair){
  !is_same_pt(names(lTCRClones[thisPair]))
  str_detect(names(lTCRClones)[thisPair], "M|LN|T") &
    str_detect(names(lTCRClones)[thisPair], "R")
}))
```

When we consider pairs that include one region in the primary tumour and another region in a metastasis, lymph node or thrombus, we find 356 TCR clones in 50 / 73 different primary metastasis pairs that share at least one TCR clone. 107 TCR clones shared in 40 / 5982 primary metastasis pairs from different patients. 

```{r}
names(normalsTCRClonotypes$data) <- paste0(str_sub(names(normalsTCRClonotypes$data), 3, 6), "-N")
names(normalsBCRClonotypes$data) <- paste0(str_sub(names(normalsBCRClonotypes$data), 3, 6), "-N")
tumNormal.TCRClonotypes <- list(data = c(TCRClonotypes$data[str_sub(names(TCRClonotypes$data), 1, 4) %in% str_sub(names(normalsTCRClonotypes$data), 1, 4)],
                                         normalsTCRClonotypes$data))
tumNormal.BCRClonotypes <- list(data = c(BCRClonotypes$data[str_sub(names(BCRClonotypes$data), 1, 4) %in% str_sub(names(normalsBCRClonotypes$data), 1, 4)],
                                         normalsBCRClonotypes$data))
```


```{r}
tumNormal.lTCRClones <- classify_clones(names(tumNormal.TCRClonotypes$data), tumNormal.TCRClonotypes)
```

```{r}
unspecificTCRSharedNormalTumour <- c()
nUnspecificTCRSharedNormalTumour <- c()
numberUnspecificPairsNormalTumour <- 0
specificTCRSharedNormalTumour <- c()
nSpecificTCRSharedNormalTumour <- c()
numberSpecificPairsNormalTumour <- 0
unspecificTCRSharedTumourTumour <- c()
nUnspecificTCRSharedTumourTumour <- c()
numberUnspecificPairsTumourTumour <- 0
specificTCRSharedTumourTumour <- c()
nSpecificTCRSharedTumourTumour <- c()
numberSpecificPairsTumourTumour <- 0

for (thisPair in 1:length(tumNormal.lTCRClones)){
  if (str_detect(names(tumNormal.lTCRClones)[thisPair], "M|LN|T|R") & str_detect(names(tumNormal.lTCRClones)[thisPair], "-N")){
    if (nrow(tumNormal.lTCRClones[[thisPair]]$shared_clones)>0){
      if(! is_same_pt(names(tumNormal.lTCRClones)[thisPair])){
        unspecificTCRSharedNormalTumour <- c(unspecificTCRSharedNormalTumour, tumNormal.lTCRClones[[thisPair]]$shared_clones$CDR3.nt)
        nUnspecificTCRSharedNormalTumour <- c(nUnspecificTCRSharedNormalTumour, length(tumNormal.lTCRClones[[thisPair]]$shared_clones$CDR3.nt))
        numberUnspecificPairsNormalTumour <- numberUnspecificPairsNormalTumour + 1
      } else {
        specificTCRSharedNormalTumour <- c(specificTCRSharedNormalTumour, tumNormal.lTCRClones[[thisPair]]$shared_clones$CDR3.nt)
        nSpecificTCRSharedNormalTumour <- c(nSpecificTCRSharedNormalTumour, length(tumNormal.lTCRClones[[thisPair]]$shared_clones$CDR3.nt))
        numberSpecificPairsNormalTumour <- numberSpecificPairsNormalTumour + 1
      }
    } 
  } 
  else 
  {
    if (nrow(tumNormal.lTCRClones[[thisPair]]$shared_clones)>0){
      if(! is_same_pt(names(tumNormal.lTCRClones)[thisPair])){
        unspecificTCRSharedTumourTumour <- c(unspecificTCRSharedTumourTumour, tumNormal.lTCRClones[[thisPair]]$shared_clones$CDR3.nt)
        nUnspecificTCRSharedTumourTumour <- c(nUnspecificTCRSharedTumourTumour, length(tumNormal.lTCRClones[[thisPair]]$shared_clones$CDR3.nt))
        numberUnspecificPairsTumourTumour <- numberUnspecificPairsTumourTumour + 1
      } else {
        specificTCRSharedTumourTumour <- c(specificTCRSharedTumourTumour, tumNormal.lTCRClones[[thisPair]]$shared_clones$CDR3.nt)
        nSpecificTCRSharedTumourTumour <- c(nSpecificTCRSharedTumourTumour, length(tumNormal.lTCRClones[[thisPair]]$shared_clones$CDR3.nt))
        numberSpecificPairsTumourTumour <- numberSpecificPairsTumourTumour + 1
      }
    }
  }
}

nPairsNormalTumSamePat <- sum(unlist(lapply(1:length(tumNormal.lTCRClones), function(thisPair) is_same_pt(names(tumNormal.lTCRClones)[thisPair]) & str_detect(names(tumNormal.lTCRClones)[thisPair], "M|LN|T|R") & str_detect(names(tumNormal.lTCRClones)[thisPair], "-N"))))
nPairsNormalTumDiffPat <- sum(unlist(lapply(1:length(tumNormal.lTCRClones), function(thisPair) ! is_same_pt(names(tumNormal.lTCRClones)[thisPair]) & str_detect(names(tumNormal.lTCRClones)[thisPair], "M|LN|T|R") & str_detect(names(tumNormal.lTCRClones)[thisPair], "-N"))))
nPairsTumTumDiffPat <- sum(unlist(lapply(1:length(tumNormal.lTCRClones), function(thisPair) ! is_same_pt(names(tumNormal.lTCRClones)[thisPair]) & str_detect(names(tumNormal.lTCRClones)[thisPair], "M|LN|T|R") & ! str_detect(names(tumNormal.lTCRClones)[thisPair], "-N"))))
nPairsTumTumSamePat <- sum(unlist(lapply(1:length(tumNormal.lTCRClones), function(thisPair) is_same_pt(names(tumNormal.lTCRClones)[thisPair]) & str_detect(names(tumNormal.lTCRClones)[thisPair], "M|LN|T|R") & ! str_detect(names(tumNormal.lTCRClones)[thisPair], "-N"))))
```

```{r}
nSpecificPairsNormalPrim_tcr = numberSpecificPairsNormalTumour
nPrimNorm_same_pat_pairs = nPairsNormalTumSamePat
nUnspecificPairsNormalPrim_tcr = numberUnspecificPairsNormalTumour
nPrimMet_diff_pat_pairs = nPairsNormalTumDiffPat
```

```{r}
tab = matrix(c(nSpecificPairsNormalPrim_tcr, nPrimNorm_same_pat_pairs, 
         nSpecificPairsPrimPrim_tcr, nPrimPrim_same_pat_pairs,
         nSpecificPairsPrimMet_tcr, nPrimMet_same_pat_pairs), nrow = 2, byrow = F)

tab2 = tab
tab2[,2] = tab2[,2] - tab[,1]
fisher.test(tab2[,1:2]) # p-value shared TCR primary vs normal = 0.01
fisher.test(tab2[,2:3]) # p-value shared TCR primary vs met = 0.29
```


```{r}
samePatientProps <- data.frame(match_TCRs = c("At least 1", "No", "At least 1", "No", "At least 1", "No"), 
                               pair_relationship = c("Primary - Normal", "Primary - Normal","Primary - Primary", "Primary - Primary", "Primary - Metastasis", "Primary - Metastasis"), 
                               prop_pairs = c(nSpecificPairsNormalPrim_tcr / nPrimNorm_same_pat_pairs, 
                                              (nPrimNorm_same_pat_pairs - nSpecificPairsNormalPrim_tcr) / nPrimNorm_same_pat_pairs, 
                                              nSpecificPairsPrimPrim_tcr / nPrimPrim_same_pat_pairs, 
                                              (nPrimPrim_same_pat_pairs - nSpecificPairsPrimPrim_tcr) / nPrimPrim_same_pat_pairs, 
                                              nSpecificPairsPrimMet_tcr / nPrimMet_same_pat_pairs, 
                                              (nPrimMet_same_pat_pairs -nSpecificPairsPrimMet_tcr)/nPrimMet_same_pat_pairs))

p = ggplot(samePatientProps, aes(x = pair_relationship, y = prop_pairs, fill = match_TCRs))  +
    geom_col() + 
    scale_fill_manual(values = c("At least 1" = "#ADD8E6", "No" = "grey80")) +
    labs(y = "Proportion", x = "") +
  scale_x_discrete(limits = c("Primary - Primary", "Primary - Metastasis", "Primary - Normal")) + 
    theme(axis.text.x = element_text(hjust = 0.5, vjust = 1, size = 5),
          axis.text.y = element_text(size = 5),
          legend.position = "none")

save_ggplot(p, file.path(FIG_DIR, "Fig6C_shared_tcr_prop"), w = 45, h = 35)
```

```{r clonality_shared_unique}
allTCRClones <- data.table::rbindlist(TCRClonotypes$data)
allTCRClones <- allTCRClones %>% as.data.frame() %>% mutate(Type = case_when(
  CDR3.nt %in% specificTCRShared ~ "shared",
  TRUE ~ "unique"
))

p = ggplot(allTCRClones, aes(x = Type, y = Proportion)) + 
  geom_violin() +
  geom_boxplot(alpha = 0.3, 
               width = 0.1, outlier.shape = NA) + 
    stat_summary(fun.y = mean, geom = "point", aes(ymax = ..y.., ymin = ..y..), col = "red", width = 0.5) + coord_flex_cart(bottom = brackets_horizontal(), left = capped_vertical("both")) + 
  ylim(c(0,0.05)) + 
  labs(x = "TCR shared in over 1 region") + 
  ggpubr::stat_compare_means(size = 2)

save_ggplot(p, file.path(FIG_DIR, "SupFig22_tcr_clonality_shared"), 
            w = 45, h = 45)
allTCRClones %>% group_by(Type) %>% summarise(mean(Proportion))
```

### BCR
```{r BCR}
# compute the total number of matching clones
sharedBCRMat <- immunarch::repOverlap(BCRClonotypes$data, .method = "public")
# arrange samples of the same patient together
sharedBCRMat <- sharedBCRMat[sort(rownames(sharedBCRMat)), sort(colnames(sharedBCRMat))]
png(filename = file.path(FIG_DIR, "sharedBCR_heatmap.png"), width = 80, height = 80, units = "mm", res = 400)
pheatmap(sharedBCRMat, cluster_rows = FALSE, cluster_cols = FALSE, show_colnames = FALSE, show_rownames = FALSE, color = hcl.colors(4, "Blues"), breaks = c(0, 1, 15, max(sharedBCRMat, na.rm = TRUE)), legend = FALSE)
dev.off()
```

We can clearly see that the BCR clones that have the same sequence are usually found in the same patient. Let's corroborate this observation

```{r}
lBCRClones <- classify_clones(names(BCRClonotypes$data), BCRClonotypes)
unspecificBCRShared <- c()
numberUnspecificPairs <- 0
specificBCRShared <- c()
numberSpecificPairs <- 0

for (thisPair in 1:length(lBCRClones)){
  if (str_detect(names(lBCRClones)[thisPair], "M|LN|T")){
    next()
  }
  if (nrow(lBCRClones[[thisPair]]$shared_clones)>0){
    if(! is_same_pt(names(lBCRClones)[thisPair])){
      unspecificBCRShared <- c(unspecificBCRShared, lBCRClones[[thisPair]]$shared_clones$CDR3.nt)
      numberUnspecificPairs <- numberUnspecificPairs + 1
    } else {
      specificBCRShared <- c(specificBCRShared, lBCRClones[[thisPair]]$shared_clones$CDR3.nt)
      numberSpecificPairs <- numberSpecificPairs + 1
    }
  }
}
```

```{r}
nSpecificPairsPrimPrim_bcr = numberSpecificPairs
nPrimPrim_same_pat_pairs = sum(sapply(1:length(lBCRClones), function (thisPair){
  is_same_pt(names(lBCRClones[thisPair])) & 
    !str_detect(names(lBCRClones)[thisPair], "M|LN|T")
}))

nUnspecificPairsPrimPrim_bcr = numberUnspecificPairs
nPrimPrim_diff_pat_pairs = sum(sapply(1:length(lBCRClones), function (thisPair){
  !is_same_pt(names(lBCRClones[thisPair])) & 
    !str_detect(names(lBCRClones)[thisPair], "M|LN|T")
}))
```

1063 different BCR clones overlapping in 240 / 262 pairs of primary sample pairs from the same patient that share at least one BCR clone. 115 different BCR clones overlapping in 14419 / 19838 pairs of primary sample pairs that share at least one BCR clone. 


```{r}
unspecificBCRSharedMets <- c()
numberUnspecificPairsMets <- 0
specificBCRSharedMets <- c()
numberSpecificPairsMets <- 0
props = data.frame(s1 = c(), s2 = c(), prop_s1 = c(), prop_s2 = c(), clone = c())
for (thisPair in 1:length(lBCRClones)){
  if (str_detect(names(lBCRClones)[thisPair], "M|LN|T") & str_detect(names(lBCRClones)[thisPair], "R")){
    if (nrow(lBCRClones[[thisPair]]$shared_clones)>0){
      if(! is_same_pt(names(lBCRClones)[thisPair])){
        unspecificBCRSharedMets <- c(unspecificBCRSharedMets, lBCRClones[[thisPair]]$shared_clones$CDR3.nt)
        numberUnspecificPairsMets <- numberUnspecificPairsMets + 1
      } else {
        specificBCRSharedMets <- c(specificBCRSharedMets, lBCRClones[[thisPair]]$shared_clones$CDR3.nt)
        df = lBCRClones[[thisPair]]$shared_clones
        s1 = colnames(df)[ncol(df) - 1]
        s2 = colnames(df)[ncol(df)]
        prop_s1 = as_vector(df[,ncol(df) - 1])
        prop_s2 = as_vector(df[,ncol(df)])
        clone = df$CDR3.nt
        props = rbind(props, data.frame(s1 = s1, s2 = s2, prop_s1 = prop_s1, prop_s2 = prop_s2, clone = clone))
        numberSpecificPairsMets <- numberSpecificPairsMets + 1
      }
    }
  }
}
```


```{r}
nSpecificPairsPrimMet_bcr = numberSpecificPairsMets
nPrimMet_same_pat_pairs = sum(sapply(1:length(lBCRClones), function (thisPair){
  is_same_pt(names(lBCRClones[thisPair])) & 
    str_detect(names(lBCRClones)[thisPair], "M|LN|T") &
    str_detect(names(lBCRClones)[thisPair], "R")
}))

nUnspecificPairsPrimMet_bcr = numberUnspecificPairsMets
nPrimMet_diff_pat_pairs = sum(sapply(1:length(lBCRClones), function (thisPair){
  !is_same_pt(names(lBCRClones[thisPair]))
  str_detect(names(lBCRClones)[thisPair], "M|LN|T") &
    str_detect(names(lBCRClones)[thisPair], "R")
}))
```

When we consider pairs that include one region in the primary tumour and another region in a metastasis, lymph node or thrombus, we find 356 BCR clones in 61 / 73 different primary metastasis pairs that share at least one BCR clone. 107 BCR clones shared in 4062 / 5982 primary metastasis pairs from different patients. 


```{r}
tumNormal.lBCRClones <- classify_clones(names(tumNormal.BCRClonotypes$data), tumNormal.BCRClonotypes)
```

```{r}
unspecificBCRSharedNormalTumour <- c()
nUnspecificBCRSharedNormalTumour <- c()
numberUnspecificPairsNormalTumour <- 0
specificBCRSharedNormalTumour <- c()
nSpecificBCRSharedNormalTumour <- c()
numberSpecificPairsNormalTumour <- 0
unspecificBCRSharedTumourTumour <- c()
nUnspecificBCRSharedTumourTumour <- c()
numberUnspecificPairsTumourTumour <- 0
specificBCRSharedTumourTumour <- c()
nSpecificBCRSharedTumourTumour <- c()
numberSpecificPairsTumourTumour <- 0

for (thisPair in 1:length(tumNormal.lBCRClones)){
  if (str_detect(names(tumNormal.lBCRClones)[thisPair], "M|LN|T|R") & str_detect(names(tumNormal.lBCRClones)[thisPair], "-N")){
    if (nrow(tumNormal.lBCRClones[[thisPair]]$shared_clones)>0){
      if(! is_same_pt(names(tumNormal.lBCRClones)[thisPair])){
        unspecificBCRSharedNormalTumour <- c(unspecificBCRSharedNormalTumour, tumNormal.lBCRClones[[thisPair]]$shared_clones$CDR3.nt)
        nUnspecificBCRSharedNormalTumour <- c(nUnspecificBCRSharedNormalTumour, length(tumNormal.lBCRClones[[thisPair]]$shared_clones$CDR3.nt))
        numberUnspecificPairsNormalTumour <- numberUnspecificPairsNormalTumour + 1
      } else {
        specificBCRSharedNormalTumour <- c(specificBCRSharedNormalTumour, tumNormal.lBCRClones[[thisPair]]$shared_clones$CDR3.nt)
        nSpecificBCRSharedNormalTumour <- c(nSpecificBCRSharedNormalTumour, length(tumNormal.lBCRClones[[thisPair]]$shared_clones$CDR3.nt))
        numberSpecificPairsNormalTumour <- numberSpecificPairsNormalTumour + 1
      }
    } 
  } 
  else 
  {
    if (nrow(tumNormal.lBCRClones[[thisPair]]$shared_clones)>0){
      if(! is_same_pt(names(tumNormal.lBCRClones)[thisPair])){
        unspecificBCRSharedTumourTumour <- c(unspecificBCRSharedTumourTumour, tumNormal.lBCRClones[[thisPair]]$shared_clones$CDR3.nt)
        nUnspecificBCRSharedTumourTumour <- c(nUnspecificBCRSharedTumourTumour, length(tumNormal.lBCRClones[[thisPair]]$shared_clones$CDR3.nt))
        numberUnspecificPairsTumourTumour <- numberUnspecificPairsTumourTumour + 1
      } else {
        specificBCRSharedTumourTumour <- c(specificBCRSharedTumourTumour, tumNormal.lBCRClones[[thisPair]]$shared_clones$CDR3.nt)
        nSpecificBCRSharedTumourTumour <- c(nSpecificBCRSharedTumourTumour, length(tumNormal.lBCRClones[[thisPair]]$shared_clones$CDR3.nt))
        numberSpecificPairsTumourTumour <- numberSpecificPairsTumourTumour + 1
      }
    }
  }
}

nPairsNormalTumSamePat <- sum(unlist(lapply(1:length(tumNormal.lBCRClones), function(thisPair) is_same_pt(names(tumNormal.lBCRClones)[thisPair]) & str_detect(names(tumNormal.lBCRClones)[thisPair], "M|LN|T|R") & str_detect(names(tumNormal.lBCRClones)[thisPair], "-N"))))
nPairsNormalTumDiffPat <- sum(unlist(lapply(1:length(tumNormal.lBCRClones), function(thisPair) ! is_same_pt(names(tumNormal.lBCRClones)[thisPair]) & str_detect(names(tumNormal.lBCRClones)[thisPair], "M|LN|T|R") & str_detect(names(tumNormal.lBCRClones)[thisPair], "-N"))))
nPairsTumTumDiffPat <- sum(unlist(lapply(1:length(tumNormal.lBCRClones), function(thisPair) ! is_same_pt(names(tumNormal.lBCRClones)[thisPair]) & str_detect(names(tumNormal.lBCRClones)[thisPair], "M|LN|T|R") & ! str_detect(names(tumNormal.lBCRClones)[thisPair], "-N"))))
nPairsTumTumSamePat <- sum(unlist(lapply(1:length(tumNormal.lBCRClones), function(thisPair) is_same_pt(names(tumNormal.lBCRClones)[thisPair]) & str_detect(names(tumNormal.lBCRClones)[thisPair], "M|LN|T|R") & ! str_detect(names(tumNormal.lBCRClones)[thisPair], "-N"))))
```

```{r}
nSpecificPairsNormalPrim_bcr = numberSpecificPairsNormalTumour
nPrimNorm_same_pat_pairs = nPairsNormalTumSamePat
nUnspecificPairsNormalPrim_bcr = numberUnspecificPairsNormalTumour
nPrimMet_diff_pat_pairs = nPairsNormalTumDiffPat
```

```{r}
tab = matrix(c(nSpecificPairsNormalPrim_bcr, nPrimNorm_same_pat_pairs, 
               nSpecificPairsPrimPrim_bcr, nPrimPrim_same_pat_pairs,
               nSpecificPairsPrimMet_bcr, nPrimMet_same_pat_pairs), nrow = 2, byrow = F)

tab2 = tab
tab2[,2] = tab2[,2] - tab[,1]
fisher.test(tab2[,1:2]) # p-value shared BCR primary vs normal = 0.83
fisher.test(tab2[,2:3]) # p-value shared BCR primary vs met = 0.62
```


```{r}
samePatientProps <- data.frame(match_BCRs = c("At least 1", "No", "At least 1", "No", "At least 1", "No"), 
                               pair_relationship = c("Primary - Normal", "Primary - Normal","Primary - Primary", "Primary - Primary", "Primary - Metastasis", "Primary - Metastasis"), 
                               prop_pairs = c(nSpecificPairsNormalPrim_bcr / nPrimNorm_same_pat_pairs, 
                                              (nPrimNorm_same_pat_pairs - nSpecificPairsNormalPrim_bcr) / nPrimNorm_same_pat_pairs, 
                                              nSpecificPairsPrimPrim_bcr / nPrimPrim_same_pat_pairs, 
                                              (nPrimPrim_same_pat_pairs - nSpecificPairsPrimPrim_bcr) / nPrimPrim_same_pat_pairs, 
                                              nSpecificPairsPrimMet_bcr / nPrimMet_same_pat_pairs, 
                                              (nPrimMet_same_pat_pairs -nSpecificPairsPrimMet_bcr)/nPrimMet_same_pat_pairs))

p = ggplot(samePatientProps, aes(x = pair_relationship, y = prop_pairs, fill = match_BCRs))  +
  geom_col() + 
  scale_fill_manual(values = c("At least 1" = "#ADD8E6", "No" = "grey80")) +
  labs(y = "Proportion", x = "") +
  scale_x_discrete(limits = c("Primary - Primary", "Primary - Metastasis", "Primary - Normal")) + 
  theme(axis.text.x = element_text(hjust = 0.5, vjust = 1, size = 5),
        axis.text.y = element_text(size = 5),
        legend.position = "none")

save_ggplot(p, file.path(FIG_DIR, "SupFigX_shared_bcr_prop"), w = 45, h = 35)
```